
\chapter{Future Scope}

MetaGPT opens several avenues for future enhancement and research. While the system already supports a range of project types---including web applications, APIs, CLIs, and libraries---one promising direction is improving the reliability and live-preview support for non-web project types such as mobile applications, microservices, and data engineering pipelines. The current E2B sandbox preview is optimised for frontend web projects; extending scaffolding logic and sandbox templates to cover additional frameworks and deployment targets would broaden the system's practical utility.

Another important area is deeper integration with developer workflows, including Git-based version control and continuous integration (CI) pipelines. Features such as automatic commit generation and pull request creation would further reduce friction between generated projects and real-world engineering practices. Additionally, incorporating richer feedback loops---for example, using runtime metrics, static analysis tools, or human-in-the-loop annotations---could help the agents iteratively improve code quality and performance over multiple pipeline runs.

From a research perspective, MetaGPT can serve as a testbed for studying collaboration between multiple LLM agents, long-lived project memory across sessions, and adaptive SOPs that evolve over time. The system currently uses Google Gemini exclusively via LangChain; experimenting with alternative LLM backends and multi-model setups, as well as hybrid workflows where human experts supervise or override specific agent stages, are also valuable directions for future work.
