%%%%%%%%%%%%%%%%%%%% BIBLIOGRAPHY %%%%%%%%%%%%%%%%%%%%
\cleardoublepage
\renewcommand\bibname{References}

\addcontentsline{toc}{chapter}{References}
\begin{thebibliography}{99}

% References for MetaGPT: A Multi-Agent Collaborative Framework with RAG Enhancement

\bibitem{ref1}
H. Hong, Y. Du, J. Pan, Z. Wu, and W. Chen, "MetaGPT: Meta Programming for Multi-Agent Collaborative Framework,"
\textit{arXiv preprint arXiv:2308.00352}, 2023.

\bibitem{ref2}
M. Chen, J. Tworek, H. Jun, Q. Yuan, H. P. de Oliveira Pinto, J. Kaplan, H. Edwards, Y. Burda, N. Joseph, G. Brockman, A. Ray, R. Puri, G. Krueger, M. Petrov, H. Khlaaf, G. Sastry, P. Mishkin, S. Chan, A. Gray, N. Ryder, M. Pavlov, A. Power, L. Kaiser, M. Bavarian, C. Winter, P. Tillet, F. Such, K. Cummings, M. Plappert, F. Chantzis, E. Barnes, A. Herbert-Voss, W. Guss, A. Nichol, A. Paino, N. Tezak, J. Tang, I. Babuschkin, S. Balaji, S. Jain, A. Saunders, C. Hesse, A. N. Carr, J. Leike, J. Achiam, V. Misra, E. Morikawa, A. Radford, M. Knight, M. Brundage, M. Murati, K. Mayer, P. Welinder, B. McGrew, D. Amodei, S. McCandlish, I. Sutskever, and W. Zaremba, "Evaluating Large Language Models Trained on Code,"
\textit{arXiv preprint arXiv:2107.03374}, 2021.

\bibitem{ref3}
Y. Li, D. Choi, J. Chung, N. Kushman, J. Schrittwieser, R. Leblond, T. Eccles, J. Keeling, F. Gimeno, A. D. Lago, J. Hubert, P. Choy, C. de Masson d'Autume, I. Babuschkin, X. Chen, P. S. Huang, J. Welbl, S. Gowal, A. Cherepanov, J. Molloy, D. J. Mankowitz, E. S. G. Robson, P. Kohli, N. de Freitas, K. Kavukcuoglu, and O. Vinyals, "Competition-Level Code Generation with AlphaCode,"
\textit{Science}, vol. 378, no. 6624, pp. 1092--1097, 2022.

\bibitem{ref4}
H. Yang, S. Yue, and Y. He, "Auto-GPT for Online Decision Making: Benchmarking and Improving,"
\textit{arXiv preprint arXiv:2304.12411}, 2023.

\bibitem{ref5}
Y. Li, D. Choi, J. Chung, N. Kushman, J. Schrittwieser, R. Leblond, T. Eccles, J. Keeling, F. Gimeno, A. D. Lago, J. Hubert, P. Choy, C. de Masson d'Autume, I. Babuschkin, X. Chen, P. S. Huang, J. Welbl, S. Gowal, A. Cherepanov, J. Molloy, D. J. Mankowitz, E. S. G. Robson, P. Kohli, N. de Freitas, K. Kavukcuoglu, and O. Vinyals, "MapCoder: Multi-Agent Code Generation for Competitive Problem Solving,"
\textit{arXiv preprint arXiv:2401.14165}, 2024.

\bibitem{ref6}
S. Yao, J. Zhao, D. Yu, N. Du, I. Shafran, K. Narasimhan, and Y. Cao, "ReAct: Synergizing Reasoning and Acting in Language Models,"
\textit{arXiv preprint arXiv:2210.03629}, 2022.

\bibitem{ref7}
N. Shinn, F. Labash, and Y. Goksel, "Reflexion: An Autonomous Agent with Dynamic Memory and Self-Reflection,"
\textit{arXiv preprint arXiv:2303.11366}, 2023.

\bibitem{ref8}
T. Schick, J. Dwivedi-Yu, R. Dessì, R. Raileanu, M. Lomeli, L. Zettlemoyer, N. Cancedda, and T. Scialom, "Toolformer: Language Models Can Teach Themselves to Use Tools,"
\textit{arXiv preprint arXiv:2302.04761}, 2023.

\bibitem{ref9}
C. Qian, X. Cong, C. Yang, W. Chen, Y. Su, J. Xu, Z. Liu, and M. Sun, "ChatDev: Communicative Agents for Software Development,"
\textit{arXiv preprint arXiv:2307.07924}, 2023.

\bibitem{ref10}
P. Lewis, E. Perez, A. Piktus, F. Petroni, V. Karpukhin, N. Goyal, H. Küttler, M. Lewis, W. Yih, T. Rocktäschel, S. Riedel, and D. Kiela, "Retrieval-Augmented Generation for Knowledge-Intensive NLP Tasks,"
\textit{Advances in Neural Information Processing Systems}, vol. 33, pp. 9459--9474, 2020.

\bibitem{ref11}
A. Vaswani, N. Shazeer, N. Parmar, J. Uszkoreit, L. Jones, A. N. Gomez, Ł. Kaiser, and I. Polosukhin, "Attention is All You Need,"
\textit{Advances in Neural Information Processing Systems}, vol. 30, pp. 5998--6008, 2017.

\bibitem{ref12}
J. Wei, X. Wang, D. Schuurmans, M. Bosma, B. Ichter, F. Xia, E. Chi, Q. Le, and D. Zhou, "Chain-of-Thought Prompting Elicits Reasoning in Large Language Models,"
\textit{Advances in Neural Information Processing Systems}, vol. 35, pp. 24824--24837, 2022.

\bibitem{ref13}
LangChain, "LangChain Documentation,"
Available: \url{https://python.langchain.com/}, Accessed: Oct. 2025.

\bibitem{ref14}
LangGraph, "LangGraph Documentation,"
Available: \url{https://langchain-ai.github.io/langgraph/}, Accessed: Oct. 2025.

\bibitem{ref15}
Google, "Gemini API Documentation,"
Available: \url{https://ai.google.dev/docs}, Accessed: Oct. 2025.

\bibitem{ref16}
Tree-sitter, "Tree-sitter Documentation,"
Available: \url{https://tree-sitter.github.io/}, Accessed: Oct. 2025.

\end{thebibliography}